flowers-pilot-analysis

### 2021-07-22 code/aa-flowers-pilot_analysis.Rmd
add data from technical pilots, full compet run, collab pilot

data_target = "technical_pilots"
##Data import constants
data_location=paste0("data/",data_target)
date_start=lubridate::ymd('2021-06-30')
image_location=paste0("figs/",data_target)

@@

#try pivot_longer instead of gather??
d.round_results <- d.round_results.raw  %>% 
  select(-chat, -context) %>%
  mutate_at(vars(starts_with('player')), as.character) %>%
  pivot_longer(cols = starts_with('player'), 
               names_to = c('blah', 'playerId', 'info'), 
               values_to = "value",
               names_sep = "_",
               values_drop_na = TRUE) %>%
  pivot_wider(names_from = "info", values_from = "value") %>%
  select(-blah) %>%
  mutate_at(c("time", "utility"), as.numeric) %>%
  #TODO: some times are negative?? interesting
  mutate(time = abs(as.numeric(time)/1000)) %>%
  rename(time_sec = time, playerResponse = response, playerUtility = utility) %>%
  arrange(createdAt, gameId, blockNum, repNum)

#exclude blocks with <12 completed rounds
rounds_exclude <- d.round_results %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=36) %>% select(gameId,blockNum)
d.round_results.final <-  d.round_results %>% anti_join(rounds_exclude)
```
```{r}
summary <- d.round_results %>% 
  group_by(gameId) %>% 
  summarize(num_rounds=max(repNum+1)) %>% 
  left_join(d.games%>% ungroup() %>% select(gameId, condition, chat=chatEnabled, gameLength))
#knitr::kable(summary)
message("Full games")
summary %>% filter(num_rounds==12)
message("Partial games")
summary %>% filter(num_rounds!=12) 
```

@@

#extract the alternative flowers
d.contexts <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  mutate(data.context = map(data.context, 
                            .f = ParseJSONColumn)) %>% 
  select("_id":"data.numPlayers") %>%
  mutate(data.context = map(data.context,
                            function(df) df %>% gather(key, value) %>%
                              mutate(key = gsub('.', '-', key, fixed=TRUE)) %>%
                              separate(key, into = c("label", "key"), sep = "-") %>%
                              spread(key, value))) %>%
  unnest(data.context) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE))
```
How often do participants select a flower that is hidden from them?

@@

```{r}
d.round_results %>% left_join(d.games %>% select(gameId, condition, chatEnabled)) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         full_condition = paste0(condition, "-", chat)) %>%
  ggplot(aes(x=log(time_sec), y = as.numeric(playerUtility), color = full_condition)) + 
  facet_grid(rows = vars(condition), cols = vars(chat)) +
  geom_point(alpha = .25) + 
  labs(x = "log time (in seconds)",
       y = "individual utility",
       title = "Response time in sec (log) and utility", subtitle =  "each participant, each round")
```

@@

```{r}
d.round_results %>% left_join(d.games %>% select(gameId, condition, chatEnabled, gameLength)) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         full_condition = paste0(condition, "-", chat)) %>%
  group_by(gameId, condition, chat, full_condition,gameLength) %>%
  summarize(totalUtility = sum(playerUtility))%>%
  ggplot(aes(x=gameLength, y = as.numeric(totalUtility), color = full_condition)) + 
  facet_grid(rows = vars(condition), cols = vars(chat), scales = "free_y") +
  geom_point(alpha = .75) + 
  labs(x = "Game length (in minutes)",
       y = "total points by group",
       title = "Length of game in mins and utility", subtitle =  "per game")

d.round_results %>% left_join(d.games %>% select(gameId, condition, chatEnabled, gameLength)) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         full_condition = paste0(condition, "-", chat)) %>%
  group_by(gameId, condition, chat, full_condition,gameLength) %>%
  summarize(totalUtility = sum(playerUtility))%>%
  mutate(totalUtility = ifelse(condition=="coopCartel", totalUtility/3, totalUtility)) %>%
  ggplot(aes(x=log(as.numeric(gameLength, units="mins")), y = as.numeric(totalUtility), color = full_condition)) + 
  #facet_grid(rows = vars(condition), cols = vars(chat), scales = "free_y") +
  geom_point(alpha = .75) + 
  geom_smooth(method=lm, alpha=.3)+
  labs(x = "Game length (in minutes)",
       y = "total points by group",
       title = "Length of game in mins and utility", subtitle =  "per game")
```

@@

Pick just one game to look at, make sure the plot is working?
```{r}
condition = "coopCartel"

d.chat %>% 
         left_join(d.games %>% select(gameId, createdAt,condition, chatEnabled)) %>% 
         filter(condition == "coopCartel",
                chatEnabled == "true") %>%
ggplot(aes(x = repNum, y=total_num_words, color = as.factor(playerId))) + 
  geom_jitter(alpha = .5) + 
  facet_grid(cols = vars(blockNum), rows = vars(gameId), scales = "free_y") + 
  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title = "Cooperative + Lang", subtitle="Number of words", y="Number of words", x="Round number", color="playerId")

d.chat %>% 
         left_join(d.games %>% select(gameId, createdAt,condition, chatEnabled)) %>% 
         filter(condition == "competCartel",
                chatEnabled == "true") %>%
ggplot(aes(x = repNum, y=total_num_words, color = as.factor(playerId))) + 
  geom_jitter(alpha = .5) + 
  facet_grid(cols = vars(blockNum), rows = vars(gameId), scales = "free_y") + 
  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title="Competitive + Lang", subtitle = "Number of words", y="Number of words", x="Round number", color="playerId")
```

@@

competcartel will have a max utility = sum(top_3_flowers)
```{r}
d.max_utility <- d.contexts %>% 
  left_join(d.games %>% select(gameId, condition)) %>%
  mutate(utility= as.numeric(utility)) %>%
  group_by(gameId, blockNum, repNum) %>% 
  slice_max(order_by =utility,n = 3)%>%
  summarize(competCartel = sum(utility),
            coopCartel = max(utility * 9)) %>% ungroup() %>%
  pivot_longer(cols = c(competCartel, coopCartel), names_to = "condition", values_to = "maxUtility")

d.utility <- d.round_results %>%
         left_join(d.games %>% select(gameId, condition, chatEnabled)) %>% 
  group_by(gameId, blockNum, repNum, trialNum,condition, chatEnabled) %>% 
  summarize(group_utility = sum(as.numeric(playerUtility)))%>% ungroup() %>%
  left_join(d.max_utility) %>%
  mutate(prop_utility = group_utility/maxUtility,
         chat = if_else(chatEnabled=="true", "Lang", "Nonlang"))

d.utility %>% 
  ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
  geom_jitter(alpha = .4) + 
  facet_grid(cols = vars(condition), rows = vars(chat), scales = "free_y") + 
  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")
```

>>>>

data_target = "collabmulti_pilot0"
##Data import constants
data_location=paste0("data/",data_target)
date_start=lubridate::ymd('2021-07-19')
image_location=paste0(here("figs",data_target))

@@

#try pivot_longer instead of gather??
d.round_results <- d.round_results.raw  %>% 
  #filter out rounds where no one made a selection
  filter(!is.na(chat)) %>%
  select(-chat, -context) %>%
  mutate_at(vars(starts_with('player')), as.character) %>%
  pivot_longer(cols = starts_with('player'), 
               names_to = c('blah', 'playerId', 'info'), 
               values_to = "value",
               names_sep = "_",
               values_drop_na = TRUE) %>%
  pivot_wider(names_from = "info", values_from = "value") %>%
  select(-blah) %>%
  mutate_at(c("time", "utility"), as.numeric) %>%
  mutate(utility = ifelse(response == "false", 0, utility)) %>%
  #TODO: some times are negative?? interesting
  mutate(time = abs(as.numeric(time)/1000)) %>%
  rename(time_sec = time, playerResponse = response, playerUtility = utility) %>%
  arrange(createdAt, gameId, blockNum, repNum)

#exclude blocks with <12 completed rounds
rounds_exclude <- d.round_results %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=36) %>% select(gameId,blockNum)
d.round_results.final <-  d.round_results %>% anti_join(rounds_exclude)
```
```{r}
game_completion <- d.round_results %>% 
  group_by(gameId) %>% 
  summarize(num_rounds=max(trialNum+1)) %>% 
  left_join(d.games%>% ungroup() %>% select(gameId, condition, chat=chatEnabled, gameLength)) %>%
  mutate(gameComplete = ifelse(num_rounds == 48, TRUE, FALSE))
#knitr::kable(summary)
message("Full games")
game_completion %>% filter(num_rounds==48)
message("Partial games")
game_completion %>% filter(num_rounds!=48) 
```

@@

#extract the alternative flowers
d.contexts <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  mutate(data.context = map(data.context, 
                            .f = ParseJSONColumn)) %>% 
  select("_id":"data.numPlayers") %>%
  mutate(data.context = map(data.context,
                            function(df) df %>% gather(key, value) %>%
                              mutate(key = gsub('.', '-', key, fixed=TRUE)) %>%
                              separate(key, into = c("label", "key"), sep = "-") %>%
                              spread(key, value))) %>%
  unnest(data.context) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE))
```

Recast games with a language USE condition
If there were fewer than 3 lines of communication, cast langUse = False
```{r}
d.games <- d.games %>% left_join(d.chat.raw %>% group_by(gameId) %>% summarize(numChat = sum(!is.na(text))) %>% 
                                   mutate(langUse = (numChat >5)) %>% select(-numChat)) %>%
  mutate(langUse = ifelse(is.na(langUse), FALSE, langUse)) %>%
#add a toddle for completed games
  left_join(game_completion %>% select(gameId, gameComplete))
```

How often do participants select a flower that is hidden from them?

@@

```{r}
d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse, gameComplete)) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         full_condition = paste0(condition, "-", chat)) %>%
  ggplot(aes(x=time_sec, y = as.numeric(playerUtility), color = full_condition)) + 
  facet_grid(rows = vars(condition), cols = vars(langUse)) +
  geom_point(alpha = .25) + 
  labs(x = "log time (in seconds)",
       y = "individual utility",
       title = "Response time in sec and utility", subtitle =  "each participant, each round")

d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse)) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         full_condition = paste0(condition, "-", chat)) %>%
  ggplot(aes(x=log(time_sec), y = as.numeric(playerUtility), color = full_condition)) + 
  facet_grid(rows = vars(condition), cols = vars(langUse)) +
  geom_point(alpha = .25) + 
  labs(x = "log time (in seconds)",
       y = "individual utility",
       title = "Response time in sec (log) and utility", subtitle =  "each participant, each round")
```

@@

```{r}
d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse, gameLength, gameComplete)) %>%
  filter(gameComplete) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         full_condition = paste0(condition, "-", chat)) %>%
  group_by(gameId, condition, chat, langUse, full_condition, gameLength) %>%
  summarize(totalUtility = sum(playerUtility))%>%
  ggplot(aes(x=gameLength, y = as.numeric(totalUtility), color = full_condition)) + 
  facet_grid(rows = vars(condition), cols = vars(langUse), scales = "free_y") +
  geom_point(alpha = .75) + 
  labs(x = "Game length (in minutes)",
       y = "total points by group",
       title = "Length of game in mins and utility", subtitle =  "per game")

d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse,gameLength, gameComplete)) %>%
  filter(gameComplete) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         chatUse = ifelse(langUse, "chatUse", "noChat"),
         full_condition = paste0(condition, "-", chatUse)) %>%
  group_by(gameId, condition, chat, full_condition,gameLength) %>%
  summarize(totalUtility = sum(playerUtility))%>%
  mutate(totalUtility = ifelse(condition=="coopCartel", totalUtility/3, totalUtility)) %>%
  ggplot(aes(x=as.numeric(gameLength, units="mins"), y = as.numeric(totalUtility), color = full_condition)) + 
  #facet_grid(rows = vars(condition), cols = vars(chat), scales = "free_y") +
  geom_point(alpha = .75) + 
  geom_smooth(method=lm, alpha=.3)+
  labs(x = "Game length (in minutes)",
       y = "total points by group",
       title = "Length of game in mins and utility", subtitle =  "per game")

ggsave(paste0(image_location, "/game_length.png"))
```

@@

Pick just one game to look at, make sure the plot is working?
```{r}
condition = "competCartel"

d.chat %>% 
         left_join(d.games %>% select(gameId, createdAt,condition, langUse, chatEnabled, gameComplete)) %>% 
         filter(langUse, gameComplete) %>%
ggplot(aes(x = repNum, y=total_num_words, color = as.factor(playerId))) + 
  geom_jitter(alpha = .5) + 
  facet_grid(cols = vars(blockNum), rows = vars(gameId), scales = "free_y") + 
  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title="Cooperative + Lang Used", subtitle = "Number of words", y="Number of words", x="Round number", color="playerId")

ggsave(paste0(image_location, "/words_over_rounds.png"))
```

@@

competcartel will have a max utility = sum(top_3_flowers)
```{r}
d.max_utility <- d.contexts %>% 
  left_join(d.games %>% select(gameId, condition)) %>%
  mutate(utility= as.numeric(utility)) %>%
  group_by(gameId, blockNum, repNum) %>% 
  slice_max(order_by =utility,n = 3)%>%
  summarize(competCartel = sum(utility),
            coopMulti= sum(utility),
            coopCartel = max(utility * 9)) %>% ungroup() %>%
  pivot_longer(cols = c(competCartel, coopMulti, coopCartel), names_to = "condition", values_to = "maxUtility")

d.utility <- d.round_results.final %>%
         left_join(d.games %>% select(gameId, condition, chatEnabled, langUse)) %>% 
  group_by(gameId, blockNum, repNum, trialNum,condition, langUse) %>% 
  summarize(group_utility = sum(as.numeric(playerUtility)))%>% ungroup() %>%
  left_join(d.max_utility) %>%
  mutate(prop_utility = group_utility/maxUtility,
         langUsed = if_else(langUse, "Lang", "Nonlang"))

d.utility %>% 
  ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
  geom_point(alpha = .4) + 
  facet_grid(cols = vars(condition), rows = vars(langUsed), scales = "free_y") + 
  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")

ggsave(paste0(image_location, "/accuracy.png"))

d.utility %>% 
  ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
  geom_point(alpha = .4) + 
  facet_grid(cols = vars(condition), rows = vars(langUsed), scales = "free_y") + 
  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")

ggsave(paste0(image_location, "/accuracy.png"))

d.utility %>% 
  filter(langUsed == "Lang") %>%
  ggplot(aes(x = repNum, y=prop_utility, color = as.factor(gameId))) + 
  geom_point(alpha = .4) + 
  facet_grid(cols = vars(blockNum), rows = vars(gameId)) + 
  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId") + ylim(0,1)

ggsave(paste0(image_location, "/accuracy.png"))
```
